

# Reproducible Research

Data Analysis 

## Load the data

```{r cache=TRUE}
data <- read.csv("activity.csv")
head(data)


```


## What is mean total number of steps taken per day?

first we take total number of steps per day
and make histogram

```{r}
num_step <- tapply(data$steps,data$date,sum,na.rm= TRUE)
hist(num_step,xlab = "Numbers of steps",main = "")
```

then find the mean median for the steps

```{r}
mean(num_step,na.rm = TRUE)

median(num_step, na.rm = TRUE)
```

## What is the average daily activity pattern?

plot of interval and average steps taken.

```{r}
avg <- aggregate(data$steps, by = list(data$interval),mean,na.rm = TRUE)
names(avg) <- c("interval","steps")
plot(avg$interval,avg$steps, type = "l",xlab = "interval",ylab = "steps",main = "Average")

```

Max number of steps in a interval

```{r}
k <- which.max(avg$steps)
avg[k,]$interval
```


## Imputing missing values

Calculating the number of missing values
```{r}
k <- is.na(data$steps)
sum(k)
```

Filling the missing value


There are many ways to fill the data 
But i am using a simple one
Inputing constant value which is mean of steps


```{r}
k <- mean(data$steps,na.rm = TRUE)
newdata <- data
newdata$steps[is.na(newdata$steps)]<- k
head(newdata)
```

histogram of average steps of new data
```{r}
num_step <- tapply(newdata$steps,newdata$date,sum,na.rm= TRUE)
hist(num_step,xlab = "Numbers of steps",main = "")
```

mean and median of new data
```{r}
mean(num_step,na.rm = TRUE)

median(num_step, na.rm = TRUE)
```

median and mean both increase as we removed na and filled with average data

## Are there differences in activity patterns between weekdays and weekends?

first we add the weekday and weekend column

```{r}
newdata$date <- as.Date(newdata$date,format = "%Y-%m-%d")
week <- sapply(newdata$date, function(d){if(weekdays(d)== "Saturday" | weekdays(d)== "Sunday"){"Weekend"}else {"Weekday"}})
newdata <- cbind(newdata,week)
head(newdata)
```


Now we plot the graph of interval and steps with weekdays


```{r}
data2 <- aggregate(steps~interval + week,newdata,mean)
library("ggplot2")
x <- ggplot(data2 , aes(x = interval,y=steps,col=week))
x+geom_line()

```



